---
title: "P8105_hw2_yx2711"
author: yx2711
date: "2022-10-03"
output: github_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```


# Problem 1

Import the `NYC_Transit_Subway_Entrance_And_Exit_Data` data set and change the variable type of route 8-11 to character type. Clean dataset name using `janitor::clean_names()`. Use `ifelse` to convert `entry` from character `YES` or `No` to logical variable type. 
The data is not tidy yet. Its dimension is 20 columns and 1868 rows.

```{r}
transit = 
  read_csv(
    "C:/Users/Carolinexyc/OneDrive/Desktop/P8105/p8105_hw2_yx2711/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

nrow(transit)
ncol(transit)
```

Uses `distinct` to obtain the number of distinct stations. There are 465 distinct stations.

```{r}
transit %>% 
  select(station_name, line) %>% 
  distinct
```

Use `filter()` to obtain the number of ADA compliant stations. There are 84 ADA compliant stations 

```{r}
transit %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Use `filter()` to obtain the proportion of station entrances / exits without vending allow entrance. There are 37.7% station station entrances / exits without vending allow entrance.

```{r}
transit %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Use `pivot_longer()` to tidy data first. Then use `filter()` and `distinct` to obtain the number of distinct station with A train and with A train and are ADA compliant.

```{r}
transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


# Problem 2

Import `Trash Wheel Collection Data` - `Mr. Trash Wheel` and clean names. 
Use `select()` to select dumpster-specific data and `as.integer()`to convert the number of sports balls to integer.
Add a variable "sheet" using `mutate()`
```{r}
mr_trash_wheel = read_xlsx("C:/Users/Carolinexyc/OneDrive/Desktop/P8105/p8105_hw2_yx2711/Trash Wheel Collection Data.xlsx", sheet = 1) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls, 0)),
    sheet = "Mr. Trash Wheel") %>% 
  drop_na(dumpster)

mr_trash_wheel

```


Import `Trash Wheel Collection Data` - `Professor Trash Wheel` and clean names. 
Use `select()` to select dumpster-specific data.
Use `mutate()` to change the year to a character type and add a variable "sheet"
```{r}
professor_trash_wheel = read_xlsx("C:/Users/Carolinexyc/OneDrive/Desktop/P8105/p8105_hw2_yx2711/Trash Wheel Collection Data.xlsx", sheet = 2) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  mutate(
    sports_balls = as.integer(NA),
    year = as.character(year),
    sheet = "Professor Trash Wheel") %>% 
  drop_na(dumpster)

professor_trash_wheel
```

Combine two datasets using `full_join()`
```{r}
trash_wheel = bind_rows(mr_trash_wheel, professor_trash_wheel)

trash_wheel

```

```{r}
nrow(trash_wheel)
ncol(trash_wheel)

weight = filter(
  trash_wheel, sheet == "Professor Trash Wheel") %>% 
  pull(weight_tons) 

sport_ball = trash_wheel %>% 
 filter(sheet == "Mr. Trash Wheel" & year == "2020") %>% 
  pull(sports_balls)

```

The resulting dataset has `r nrow(trash_wheel)` observations and `r ncol(trash_wheel)` variables.
Some examples of key variables are weight, volume, plastic bottles, polystyrene and so on. 
The total weight of trash collected by professor trash wheel is `r sum(weight)` tons.
The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(sport_ball)`.



# Problem 3

Import `pols-month` dataset. Use `separate()` and `mutate()` to break up variable month.
Use `ifelse()` to create a president variable.
Use `select()` to drop variable.
```{r}
pols_month = 
  read.csv("C:/Users/Carolinexyc/OneDrive/Desktop/P8105/p8105_hw2_yx2711/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
    day = as.integer(day),
    president = ifelse(prez_gop == 0, "dem", "gop" )
  ) %>% 
  select(-day, -prez_dem, -prez_gop)
 
```


Import `snp` dataset. Use `separate()` and `mutate()` to break up variable month. 
```{r}
snp = 
  read.csv("C:/Users/Carolinexyc/OneDrive/Desktop/P8105/p8105_hw2_yx2711/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year","month", "day")) %>% 
   mutate(
    year = as.integer(year),
    year = ifelse(year > 2015, year - 100, year),
    month = month.name[as.integer(month)],
    day = as.integer(day),
  ) %>% 
  relocate(year, month) %>% 
  select(-day)

```

Import `unemployment` dataset
Use `pivot_longer()` to convert the dataset from wide to long format.
```{r}
unemployment = 
  read.csv("C:/Users/Carolinexyc/OneDrive/Desktop/P8105/p8105_hw2_yx2711/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "Unemployed"
  ) %>% 
  mutate(
    month = str_to_title(month),
    month = month.name[match(month, month.abb)]
  )

```

Merge snp into pols
```{r}
snp_pols =
  left_join(pols_month, snp)

```

Merging unemployment into the result
```{r}
result = 
  left_join(snp_pols, unemployment)

nrow(result)
ncol(result)
max(result$year)
min(result$year)
```

The pols_month dataset contains in total `r ncol(pols_month)` variables and `r nrow(pols_month)` observations, and key variables are year, month, and president.  
The snp dataset contains `r ncol(snp)` variables and `r nrow(snp)` observations, and key variables are year, month, and close.
The unemployment dataset contains `r ncol(unemployment)` variables and `r nrow(unemployment)` observations, and key variables are year, month and unemployment.  
The resulting dataset has `r nrow(result)` rows and `r ncol(result)` columns. The range of years is from `r min(result$year)` to `r max(result$year)`. The key variables are year, month, president, close, and unemployment. 

